---
title: "Human Activity Recognition with Smartphones"
output: html_notebook
---

```{r, warning = FALSE, message=FALSE}
rm(list=ls())

library(knitr)
library(ggplot2)
library(plyr)
library(dplyr)
library(corrplot)
library(caret)
library(gridExtra)
library(scales)
library(Rmisc)
library(ggrepel)
library(randomForest)
library(psych)
library(xgboost)
library(factoextra)
library(pROC)
library(parallel)
library(doParallel)
library(kernlab)
library(MLmetrics)
```

## Read Data 

```{r}
train <- read.csv("~/Github/Human-Activity-Recognition-with-Smartphones/Data/train.csv")
test <- read.csv("~/Github/Human-Activity-Recognition-with-Smartphones/Data/test.csv")
```

### Explort the data

```{r}
dim(train)
```

```{r}
dim(test)
```

### column types

```{r}
colnames = colnames(train)
m = length(colnames(train))

coltype <- vector()
colstat <- vector()
for (j in 1:m){
  temp=strsplit(colnames[j],"[.]")[[1]]
  coltype[j] = temp[1]
  colstat[j] = temp[2]
}
print(table(coltype))
print(table(colstat))
```

### combine the train and test 

```{r}
df <- rbind(train, test)
dim(df)
```

### column data type

```{r}
table(sapply(train, class))
```

interger is subject column

### check na

```{r}
sum(is.na(df))
```

### The response variable; Activity

From below we can see, 'Activiies' can be cateogirzed into two ways:
1. static and moving
2. up, down and same level

We will use these informaion to create 

```{r,fig.width=10, fig.height=3.5}
ggplot(data=df[!is.na(df$Activity),], aes(x=Activity)) +
  geom_bar(fill="blue", stat='count', width=.75) +
  coord_cartesian(ylim = c(1250, 2000))
table(df$Activity)
```

The response varaible is balanced. We do not need to perform sampling techniques.

### 7-characteristics: angle

```{r}
colcheck <- function(df,pattern){
  colnames = colnames(df)
  df_subset <- df[,sapply(colnames,grepl,fixed=TRUE,pattern=pattern)]
  colnames_df_subset = colnames(df_subset)
  colstat <- vector()
  m = length(colnames_df_subset)
  for (j in 1:m){
    temp=strsplit(colnames_df_subset[j],"[.]")[[1]]
    colstat[j] = temp[2]
  }
  print(kable(table(colstat)))
  print(describe(df_subset))
  return(df_subset)
}

df_angle = colcheck(df,"angle")
```

```{r}
cor(df_angle)
```

### 79 characteristics: fBodyAcc

```{r}
df_fBodyAcc = colcheck(df,"fBodyAcc.")
```

#### fBodyAcc.bandsEnergy

```{r}
df_fBodyAcc.bandsEnergy = colcheck(df,"fBodyAcc.bandsEnergy")
```

### 13-characterstics fBodyBodyAccJerkMag

```{r}
df_fBodyAccJerk = colcheck(df,"fBodyBodyAccJerkMag")
```

### 40-characteristic: tBodyAcc

```{r}
df_tBodyAcc = colcheck(df,"tBodyAcc.")
```

#### tBodyAcc.arCoeff

```{r}
df_tBodyAcc.arCoeff = colcheck(df,"tBodyAcc.arCoeff")
```

#### tBodyAcc.correlation

```{r}
df_tBodyAcc.correlation = colcheck(df,"tBodyAcc.correlation")
```

## Feattures engineering

### dimension reduction

We are going to use PCA to do dimension reduction on time-serise data

arCoeff(): Autorregresion coefficients with Burg order equal to 4
use pca to 

```{r}
char_40_var_cols = c("tBodyAcc.","tBodyAccJerk.","tBodyGyro.","tBodyGyroJerk.","tGravityAcc.")
colnames = colnames(df)
for(col in char_40_var_cols){
  pattern = paste(col,"arCoef",sep="")
  df_subset = df[,sapply(colnames(df),grepl,fixed=TRUE,pattern=pattern)]
  df_subset_PCA = prcomp(df_subset)
  cumpve = summary(df_subset_PCA)$importance
  selected_pc_cols= sum(cumpve["Cumulative Proportion",] < 0.8) + 1
  selected_pcs = df_subset_PCA$x[,(1:selected_pc_cols)]
  selected_pcs_colnames = vector(length=selected_pc_cols)
  for(i in 1:selected_pc_cols){
    selected_pcs_colnames[i] = paste(pattern,paste("pc",i,sep=""),sep=".")
  }
  colnames(selected_pcs) <- selected_pcs_colnames
  df[,colnames(df_subset)] <- NULL
  df= cbind(df,selected_pcs)
}
```

79-characteristic: bandsEnergy

```{r}
char_79_var_cols = c("fBodyAcc.","fBodyAccJerk.","fBodyGyro.")
for(col in char_79_var_cols){
  pattern = paste(col,"bandsEnergy",sep="")
  df_subset = df[,sapply(colnames(df),grepl,fixed=TRUE,pattern=pattern)]
  df_subset_PCA = prcomp(df_subset)
  cumpve = summary(df_subset_PCA)$importance
  selected_pc_cols= sum(cumpve["Cumulative Proportion",] < 0.8) + 1
  selected_pcs = df_subset_PCA$x[,(1:selected_pc_cols)]
  selected_pcs_colnames = vector(length=selected_pc_cols)
  for(i in 1:selected_pc_cols){
    selected_pcs_colnames[i] = paste(pattern,paste("pc",i,sep=""),sep=".")
  }
  colnames(selected_pcs) <- selected_pcs_colnames
  df[,colnames(df_subset)] <- NULL
  df= cbind(df,selected_pcs)
}
```

### Moving and Static 

The idea is coming from the below kernel
https://www.kaggle.com/vikashrajluhaniwal/eda-all-classification-algorithms-with-96-acc

```{r,fig.width=10, fig.height=3.5}
ggplot(data=df, mapping=aes(x=Activity, y = fBodyAcc.mean...X + fBodyAcc.mean...Y))+
  geom_boxplot(outlier.shape=NA) +
  ylab("fBodyAcc.mean.X + fBodyAcc.mean.Y")
```

```{r}
df$dynamic_static = ifelse(df$fBodyAcc.mean...X+df$fBodyAcc.mean...Y > -1.5, "dynamic","static")
df$dynamic_static = as.factor(df$dynamic_static)
```


### Resting

```{r,fig.width=10, fig.height=3.5}
ggplot(data = df, mapping = aes(x=Activity, y=angle.X.gravityMean.)) +
  geom_boxplot(outlier.shape=NA)
```

```{r}
df$resting = ifelse(df$angle.X.gravityMean. > 0,1,0)
df$resting = as.factor(df$resting)
```

# upward_downward_same

```{r}
df$upward_downward_same = 0
df[df$Activity == "WALKING_UPSTAIRS",c("upward_downward_same")] = 1
# df[df$Activity == "WALKING_DOWNSTAIRS",c("upward_downward_same")] = -1
df$upward_downward_same = as.factor(df$upward_downward_same)

set.seed(2018)
quick_RF <- randomForest(x=subset(df[1:7352,],select=-c(upward_downward_same,Activity)), y=df$upward_downward_same[1:7352], ntree=100,importance=TRUE)
imp_RF <- importance(quick_RF)
imp_DF <- data.frame(Variables = row.names(imp_RF), feat_imp = imp_RF[,1])
imp_DF <- imp_DF[order(imp_DF$feat_imp, decreasing = TRUE),]
```

```{r}
ggplot(imp_DF[1:20,], aes(x=reorder(Variables, feat_imp), y=feat_imp, fill=feat_imp)) + 
  geom_bar(stat = 'identity') + labs(x = 'Variables', y= 'Mean Decrease of Accuracy') + coord_flip() + theme(legend.position="none")
```
```{r}
imp_DF
```

```{r}
ggplot(data = df, mapping = aes(x=Activity, y=-tGravityAcc.arCoef.pc1
-tGravityAcc.min...Y	
-tGravityAcc.mean...Y
#+tGravityAcc.energy...Y
#+fBodyGyro.maxInds.Z
-fBodyAcc.meanFreq...Z
+angle.Y.gravityMean.
#+fBodyGyro.kurtosis...Y
-fBodyAccJerk.maxInds.Z
+fBodyBodyAccJerkMag.max..)) + 
  geom_boxplot(outlier.shape=NA) + 
  ylab("upstairs")
```


```{r}
df$upward_downward_same = NULL
df$upstairs = with(df, -tGravityAcc.arCoef.pc1 - tGravityAcc.min...Y - tGravityAcc.mean...Y - fBodyAcc.meanFreq...Z + angle.Y.gravityMean. - fBodyAccJerk.maxInds.Z + fBodyBodyAccJerkMag.max..)
df$upstairs = ifelse(df$upstairs > 1, 1,0)
df$upstairs = as.factor(df$upstairs)
```

### Modelling

```{r}
train <- df[1:7352,]
test <- df[7353:10299,]
```

#### Random Forest

```{r}

mtryValues <- c(30,50,70,100,200)

cl <- makePSOCKcluster(8)
registerDoParallel(cl)
start_time_rf <- Sys.time()
ctrl <- trainControl(method="repeatedcv", number=10, repeats=5, search="random",classProbs = TRUE)

set.seed(476)
rf <- train(x = subset(train,select=-c(subject,Activity)), 
            y = train$Activity,
            method = "rf",
            ntree = 500,
            tuneGrid = data.frame(mtry = mtryValues),
            importance = TRUE,
            metric = "Kappa",
            trControl = ctrl)

stopCluster(cl)
end_time_rf <- Sys.time()
end_time_rf - start_time_rf

##Pro tip: save your results immediately after running (in case you have to leave your computer)
saveRDS(rf, '~/Github/Human-Activity-Recognition-with-Smartphones/RF_df.rds')
```

#### XGBoost

```{r}
set.seed(476)
registerDoParallel(cl)
xgb <- train(x=subset(train,select=-c(subject,Activity)), 
             y=train$Activity,
             method="xgbTree",
             trControl=trnCtrl,
             tuneLength=3)
saveRDS(xgb, '~/Github/Human-Activity-Recognition-with-Smartphones/xgb.rds')
stopCluster(cl)
end_time_xgb<- Sys.time()
end_time_xgb - start_time_xgb
```

#### Support Vector Machines

```{r}
set.seed(476)
registerDoParallel(cl)
ctrl <- trainControl(method = "repeatedcv",
                     summaryFunction = multiClassSummary,
                     classProbs = TRUE,
                     savePredictions = TRUE,
                     repeats = 5)

sigmaRangeFull <- sigest(as.matrix(subset(train,select=-c(subject,Activity))))
svmRGridFull <- expand.grid(sigma =  as.vector(sigmaRangeFull),
                            C = 2^(-3:3))
svm <- train(x = subset(train,select=-c(subject,Activity)),
             y = train$Activity,
             method = "svmRadial",
             metric = "ROC",
             tuneGrid = svmRGridFull,
             trControl = ctrl)
saveRDS(svm, '~/Github/Human-Activity-Recognition-with-Smartphones/svm.rds')
stopCluster(cl)
```

#### Neutral Networks

```{r}
nnetGrid <- expand.grid(size = 1:10, decay = c(0, .1, 1, 2))
maxSize <- max(nnetGrid$size)
registerDoParallel(cl)
nnetFit <- train(x = subset(train,select=-c(subject,Activity)), 
                 y = train$Activity,
                 method = "nnet",
                 metric = "ROC",
                 tuneGrid = nnetGrid,
                 trace = FALSE,
                 maxit = 2000,
                 MaxNWts = 1*(maxSize * (ncol(subset(train,select=-c(subject,Activity))) + 1) + maxSize + 1),
                 trControl = ctrl)
saveRDS(nnetFit, '~/Github/Human-Activity-Recognition-with-Smartphones/nnetFit.rds')
stopCluster(cl)
```



